### 2020-02-11 paper/paper.tex
updated bao figure

% This project is part of the Continuous-Function Estimator project
% Copyright 2019 the authors.

% to-do
% -----

% style notes
% -----------
% - line break at sentence breaks

%\documentclass[twocolumn]{aastex62}
%\documentclass[12 pt]{article}
\documentclass[modern]{aastex62}

\usepackage[sort&compress]{natbib}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage{xspace}
\usepackage{xcolor}

% aastex parameters
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
\received{XXX}
%\revised{not yet}
\accepted{YYY}
%\submitjournal{ApJ}
\shorttitle{A Continuous Correlation Function Estimator}
\shortauthors{Storey-Fisher and Hogg}

% language
\newcommand{\cf}{2pcf\xspace} %2pF? 2PCF? %TODO: fix spacing after
% to capitalize or not to capitalize? leaning no
\newcommand{\Est}{The Continuous-Function Estimator\xspace}
\newcommand{\est}{the Continuous-Function Estimator\xspace}
\newcommand{\LS}{LS\xspace}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\etc}{\foreign{etc}}
% math
\newcommand{\inv}{^{-1}}
\newcommand{\T}{^{\mathsf{T}}}
\newcommand{\hmpc}{$h^{-1}$Mpc}
% comments
\newcommand{\KSF}[1]{\textcolor{teal}{KSF says: #1}}
%\newcommand{\KSF}[1]{\textcolor{teal}{}}

% margins
%\addtolength{\topmargin}{-0.75in}
%\addtolength{\textheight}{1.50in}

% affiliations
\newcommand{\ccpp}{\affiliation{%
    Center for Cosmology and Particle Physics,
    Department of Physics,
    New York University}}
\newcommand{\flatiron}{\affiliation{%
    Flatiron Institute, Simons Foundation}}
\newcommand{\cds}{\affiliation{%
    Center for Data Science,
    New York University}}
\newcommand{\mpia}{\affiliation{%
    Max-Planck-Institut f\"{u}r Astronomie, Heidelberg}}

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing

\title{\textbf{Two-point statistics without bins: A continuous-function generalization of the correlation function estimator for large-scale structure}}
%Two-point statistics without bins: A correlation function estimator for large-scale structure in a continuous-function basis
%Two-point statistics without bins: The correlation function estimator for large-scale structure generalized to a continuous-function basis
%\title{\textbf{A Continuous-Function Estimator of the Correlation Function for Large-Scale Structure}}
%\title{\textbf{A Continuous Correlation Function Estimator for Large-Scale Structure}}
%\title{\textbf{A Continuous-Function Correlation Function Estimator for Large-Scale Structure}}
%\title{\textbf{Projecting the Correlation Function onto Continuous Functions for Large-Scale Structure Surveys}
%\title{\textbf{A Continuous Representation of the Correlation Function Estimator for Large-Scale Structure}}
%\title{\textbf{Correlation Function Estimation with Continuous Functions for Large-Scale Structure}}
%\title{\textbf{Estimating the Two-Point Correlation Function with Continuous Functions}}
%\title{\textbf{Projecting Two-Point Correlations onto General Basis Functions: A New Estimator For Large-Scale Structure}}
%\title{\textbf{A Correlation Function Estimator with General Basis Functions for Large-Scale Structure}}
%\title{\textbf{A Generalized Correlation Function Estimator For Large-Scale Structure}}
%\title{\textbf{No More Bins:\\ A Vectorized Correlation Function Estimator For Large-Scale Structure}}
%\title{A Generalized Correlation Function Estimator for Galaxy Surveys}

%Generalized Estimator?
%Projected Estimator?
%Continuous Estimator?
%Continuous-Function Estimator?
%Linear-Regression Estimator?
%Vectorized estimator?

\author[0000-0001-8764-7103]{Kate Storey-Fisher}
\ccpp

\author[0000-0003-2866-9403]{David W. Hogg}
\ccpp
\cds
\mpia
\flatiron

\begin{abstract}\noindent
% Context
The two-point correlation function (\cf) is the most important statistic in structure formation, used to measure the clustering of density field tracers (e.g. galaxies).
Current estimators of the 2pcf, including the standard Landy-Szalay (\LS) estimator, have significant limitations.
%Aims
In this work we address the issue that the \LS estimator evaluates the \cf in bins of separation between objects, which results in a loss of information and a trade-off between bias and variance. \KSF{there are other aims but they are hard to describe, as they are not ``issues'' with the LS estimator but other capabilities of ours - how to phrase?}
%Methods
We present a new estimator for the \cf, \emph{\est}, which generalizes \LS to a continuous representation and obviates binning in separation or any other property. 
Our estimator replaces the binned pair counts of \LS with a linear superposition of any set of basis functions, and outputs the best-fit linear combination of basis functions to describe the \cf. 
It is closely related to the information-theory optimal estimator used in linear least-squares fitting. The choice of basis can take into account the expected form of the \cf, as well as its dependence on other properties beyond separation.
%Results
We show that \est can estimate the clustering of artificial data in representations that provide more accuracy with fewer basis functions than \LS.
Using a spline basis representation, we show that \est achieves lower bias and lower variance than \LS. 
We also demonstrate how the estimator can be used to directly estimate the Baryon Acoustic Oscillation scale.
Critically, these representations reduce the number of mock catalogs required for covariance estimation, a limiting factor in \cf measurements.
%As a demonstration, we apply the estimator to the Luminous Red Galaxy sample of the Sloan Digital Sky Survey (SDSS). 
%We show that the continuous-function estimator, with basis functions chosen to be derivatives with respect to the cosmological parameters, provides a direct measurement of cosmological parameters from galaxy observations.
%Similarly, we use the estimator to directly measure the location of the baryon acoustic feature in the SDSS data.
We discuss other applications and limitations of \est for present and future studies of large-scale structure, including determining the dependence of clustering on galaxy properties and potentially unifying real-space and Fourier-space approaches to clustering measurements.

\end{abstract}

\keywords{cosmology: large-scale structure --- galaxies: statistics} %?

\section{Introduction}

The large-scale structure (LSS) of the universe is a critical probe of fundamental cosmology. 
It encodes information about the physics of the early universe and the subsequent expansion history.
In particular, the LSS provides a probe of the Baryon Acoustic Oscillations (BAO), density fluctuations resulting from baryon-photon coupling in the early universe.
The distance traveled by these waves imprints a feature on the statistical description of the LSS, which can be used to determine the characteristic BAO length scale \citep{EisensteinHu1998}.
The LSS also contains the signature of redshift-space distortions caused by the peculiar velocities of galaxies, which probe the growth rate of structure \citep{Kaiser1987}.
Additionally, the LSS can be used to constrain galaxy formation in conjunction with models of galaxy bias (e.g. \citealt{Hamilton1988}). %?
With current observations, the LSS is well-described by a cold dark matter model with a cosmological constant, the standard $\Lambda$CDM model.
Upcoming galaxy surveys will observe larger volumes with improved measurements, allowing us to test $\Lambda$CDM to even higher precision.
\KSF{Should I name-drop surveys, and their data volume / sky area?}

We characterize the LSS by using luminous sources to trace the underlying matter density field.
These tracers are often to taken to be galaxies, but can also be galaxy clusters, quasars and other sources.
\KSF{should i say that hereafter we will take them to be galaxies?}
The clustering of these objects is measured with two-point statistics, namely the power spectrum $P(k)$ and the two-point correlation function (\cf).
These characterize the clustering in Fourier space and real space, respectively, with the \cf defined as the Fourier Transform of the power spectrum: 
\begin{equation}
\xi(r) = \frac{1}{2\pi^2} \int_0^{\infty} P(k) \frac{\mathrm{sin}(kr)}{kr} k^2 dk
\end{equation}
where we assume isotropy. \KSF{should i give the original Fourier transform before itegrating over orientation?}
In principle, the power spectrum and the two-point correlation function contain the same information.
However, in practical applications the survey boundaries introduce nontrivial issues in computing these statistics, leading to diverging approaches to their computation with a significant difference in expense.
The \cf requires more computational power and extra survey products, but it is an incredibly useful tool; for instance, it well-suited to the analysis of the BAO feature which manifests at a single scale in real space.

The \cf measures the excess probability that any two galaxies are separated by a given distance, compared to a uniform distribution; effectively, it characterizes the strength of clustering at a given spatial scale. 
% need to discuss statistical homogeneity & isotropy? (ergodicity?)
In calculating the \cf, the boundaries of the surveys prevent us from directly summing pair counts due to nontrivial edge effects.
To account for the survey boundaries as well as corrupted regions due to issues such as bright foreground stars, a set of random points are Poisson-distributed within the acceptable survey window. 
The pairwise correlations of these unclustered points are used to normalize out the survey window when estimating the \cf of the clustered data.
Typically, this requires random points on the order of 10-100 times the number of data points, making the random correlations the limiting factor in \cf computation.

The \cf is computed in bins of radial separation, meaning that in practice it measures the volume average of the \cf over the bin. %is this accurate?
This binning introduces inherent limitations.
First, the choice of bins requires a trade-off between bias and variance: fewer bins may bias the result, while more bins increases the variance of measurement.
Finite-width bins also result in a loss of information about the property in which one is binning.
As we strive for extreme precision in large-scale structure analyses, we should be maximizing the information we extract from the data.

More generally, binning adds arbitrary boundaries between continuous data; results should not depend on bin choice, yet they sometimes do.
\cite{Lanzuisi2017} noted that the choice of binning axis impacts the detected correlation between the luminosity of active galactic nuclei and their host galaxies; \cite{Grimmett2020} devised a method to investigate this correlation in a continuous manner using a hierarchical Bayesian model, eliminating the need for binning.
\cite{Bailoni2016} explored the dependence of clustering analyses on the number of redshift bins, finding a non-negligible difference in cosmological parameter uncertainties.
The implications for BAO analyses were explored by \citep{Percival2014}, who found that there is an optimal bin width given the analysis method.
This balances the increasing statistical error with small bins and the offset in the derived BAO peak location with large bins; the effects are small but non-negligible.
%Moreover, we note that standard estimators are fundamentally biased due to this finite binning; we will address this in a future work. %TODO: figure out this wording. need to explain more than this or leave out entirely?

%some papers have equations (even this eqn) in intro - but should bump it to motivation and just explain?
Estimators of the \cf have been studied extensively (\citealt{PeeblesHauser1974}; \citealt{DavisPeebles1983}; \citealt{Hamilton1993}).
The current standard estimator was proposed by \cite{LandySzalay1993}, hereafter \LS. It is based on summing all data pairs $DD$ with a given separation and using data-random pairs $DR$ and random pairs $RR$ to correct for the survey boundary. The correlation function $\xi_k(r)$ for the $k^\mathrm{th}$ separation bin is
%assume we should keep this notation even though it's clunky, as it's recognizable
%should we use estimator notation with a hat on xi?
\begin{equation}
\xi_k(r) = \frac{DD_k(r) - 2DR_k(r) + RR_k(r)}{RR_k(r)},
\end{equation}
where we assume the pair counts are normalized. %should we write out normalization explicitly? here or somewhere else?
Compared with other estimators based on simple combinations of $DD$, $DR$ and $RR$, \LS has been shown to have the lowest bias and variance \citep{Kerscher2000}. %added simple because there was the recent paper which looked at all linear combos and optimized, think that one might have better bias/variance properties
Estimators of the \cf must also take into account the imperfect nature of the survey area that is in the window, including the target completeness and fiber collisions; typically each galaxy pair is assigned a weight based on these. %is this necessary? not sure where it fits
Then, pair counts are replaced by the sum of pair weights.

Variations on the random catalog pair count method have been proposed in recent years.
\cite{Demina2016} replaced the $DR$ and $RR$ terms with an integral over the probability map, reducing computation time and increasing precision.
An estimator proposed by \cite{VargasMagana2013} iterates over sets of mock catalogs to find an optimal linear combination of data and random pair counts, reducing the bias and variance.
The marked correlation function \citep{WhitePadmanabhan2009} sums weights based on properties of the galaxies one is interested in, such as the local density or galaxy luminosity, and avoids the use of a random catalog.
The estimators described so far have all taken probabilistic approaches; others have taken a likelihood approach.
\cite{BaxterRozo2013} introduced a maximum likelihood estimator for the \cf, which achieves lower variance compared to the \LS estimator, enabling finer binning and requiring a smaller random catalog for the same precision.

These estimators present improvements to \LS, but they are still limited to estimates in separation bins.
Some require additional computational costs or layers of complexity, so the standard formulation of \LS continues to be the default estimator used in most analyses.

In this paper, we present a new estimator for the correlation function, \est, which generalizes the \LS estimator to produce a continuous estimation of the \cf. 
\Est projects the galaxy pairs onto a set of continuous basis functions and computes the best-fit linear combination of these functions.
The basis representation can depend on the pair separation as well as other desired properties, and can also utilize the known form of the \cf.
For top-hat basis functions, \est exactly reduces to the \LS estimator. 
This estimator removes the need for binning and allows for the \cf to be represented by fewer basis functions, requiring fewer mock catalogs to compute the covariance matrix.
It is particularly well-suited to the analysis of LSS features such as the BAO peak; we find that we can accurately locate the peak with fewer components.

%will obviously keep changing this as paper develops
%spell out acronyms again that were defined in the abstract?
This paper is organized as follows. 
In $\S$\ref{sec:motiv}, we motivate our estimator and explain its formulation.
%In Section \ref{sec:est}, we prove its correctness and describe our implementation, and demonstrate its application on a simulated data set. 
%Our results on the SDSS LRG sample are shown in Section \ref{sec:app}. 
We demonstrate its application on a simulated data set, including a toy BAO analysis, in $\S$\ref{sec:app}.
We discuss the implications and other possible applications in $\S$\ref{sec:discuss}.

\section{Motivation and Formulation} 
\label{sec:motiv}

\subsection{Standard Clustering Estimation}

We assume we have a data catalog with $N_D$ objects and a random catalog with $N_R$ objects.
The pair counts for \LS and related estimators are then defined explicitly as
%TODO: fix spacing on sides of equal sign
\begin{eqnarray}\displaystyle
DD_k &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} i(g_k < |x_n - x_{n'}| < h_k) \\
DR_k &\equiv& \frac{1}{N_D N_R} \sum_{n m} i(g_k < |x_n - x_m| < h_k) \\
RR_k &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} i(g_k < |x_m - x_{m'}| < h_k),
\end{eqnarray}
where $DD_k$ is the count of data--data tracer pairs in bin $k$ (which has bin edges $g_k$ and $h_k$), $i()$ is an indicator function \KSF{define?}, $x$ is the tracer position, the $n$ and $n'$ indices index data positions, the $m$ and $m'$ indices index random catalog positions, $DR_k$ is the count of data--random pairs, and $RR_k$ is the count of random--random pairs.
\KSF{Do i need to note that Corrfunc (and some papers) actually kills the -1 in the prefactor? though this might be related to how Corrfunc does pair counting}

The \LS estimator has been analyzed extensively; it is shown to have the lowest bias and variance compared to other pair-count-based estimators (e.g. \citealt{Kerscher2000}).
In the limit of unclustered data, for a data volume much larger than the scales of interest, and an infinitely large random catalog, the \LS estimator is known to be optimal: it is unbiased and has minimum variance.
In practice the latter two limits are sufficiently satisfied, but the data we are interested in is clustered.
\LS does show a bias on very large scales ($>$130 \hmpc), but the bias is significantly smaller than that of most other estimators (\citealt{Kerscher1999}, \citealt{VargasMagana2013}).
\LS is also less sensitive to the number of random points than other estimators \citep{Kerscher2000}.
\cite{VargasMagana2013} show that for clustered data, the \LS estimator has lower variance than other estimators, but does not reach the Poisson noise limit.

\KSF{Figure demonstrating trade-off between bias and variance? or too simple}

\KSF{Should i explain / show some of the other estimators? --> now i do in the beyond LS section}

\KSF{Do i need to explain real space vs redshift space vs projected correlation function?}

\subsection{Least Squares Fitting}

Estimating clustering is closely related to least-squares fitting.
We are trying to find the best representation of spatial data in the space of two-point radial separation.
Recall that the linear least-squares fit to a set of data 
\begin{equation}
X = [A\T C\inv A]\inv [A\T C\inv Y]
\end{equation}
where $X$ is the vector of best-fit parameters, $A$ is a matrix with zeroth and first order terms of $x$ data, $C$ is the covariance matrix, and $Y$ is a column vector of $y$ data.
The second bracketed term contains the observed data; the first bracketed term weights the data by the errors.
Naively, in the case of the \cf, the observed data is the pair counts at a given separation, and the weights are provided by the pair counts of the random catalog.
Indeed, this is reminiscent of the so-called natural estimator of the \cf, $\xi_k = DD_k/RR_k - 1$ (e.g. \citealt{Kerscher2000}).

From this connection, we can infer the form of the estimator.
\KSF{expand this section}

% TODO: fix this capitalization issue
\subsection{\Est}
\label{sec:est}

We generalize the \LS estimator defined above to any set of $K$ basis functions $f$ of the pair, so we now have
\begin{eqnarray}\displaystyle
DD &\equiv& \frac{2}{N_D(N_D-1)} \sum_{n n'} f(T_n, T_{n'}) \\
DR &\equiv& \frac{1}{N_D N_R} \sum_{n m} f(T_n, T_{m}) \\
RR &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \\
QQ &\equiv& \frac{2}{N_R(N_R-1)} \sum_{m m'} f(T_m, T_{m'}) \cdot f\T(T_m, T_{m'}),
\end{eqnarray}
%introduced this T notation to be the most general but it's a bit clunky (see below)
%what do we think about K-vector notation?
%matrix->tensor?
where $DD$, $DR$, and $RR$ are now $K$-vectors, $T$ refers to the data for the given tracer, and $QQ$ is a $K$-by-$K$ matrix defined as the outer product of basis function evaluations of the random--random pairs.

Then, we can compute the \cf as
\begin{eqnarray}\displaystyle
% here is where the T notation is weird - when you look at evaluation
a &\equiv& QQ\inv \cdot (DD + 2DR - RR) \\
\xi(T_l, T_{l'}) &\equiv& a\T \cdot f(T_l, T_{l'})
\end{eqnarray}
where $T_l$ and $T_{l'}$ contain the data values at which to evaluate $\xi$, and $a$ is a $K$-vector of the ``amplitudes'' of the basis functions.
This generalized two-point estimator, or \est, removes the need for binning as the basis functions can be continuous.
It is invariant under affine transformations; we show this in Appendix \ref{sec:affine}.

%%% INDEX NOTATION VERSION %%%
\iffalse % Messier and maybe inaccurate (with the inverse QQ)? but more consistent
\begin{eqnarray}\displaystyle
DD_k &\equiv& \sum_{n n'} f_k(T_n, T_{n'}) \\
DR_k &\equiv& \sum_{n m} f_k(T_n, T_{m}) \\
RR_k &\equiv& \sum_{m m'} f_k(T_m, T_{m'}) \\
QQ_{kl} &\equiv& \sum_{m m'} f_k(T_m, T_{m'}) f_l(T_m, T_{m'}),
\end{eqnarray}
%should we change the letter of k to mean basis functions rather than bins now?
%better notation for T?
Then, we can compute the \cf as
\begin{eqnarray}\displaystyle
\xi_k(T_j, T_{j'}) &\equiv& a_k f_k(T_j, T_{j'}) \\
a_k &\equiv& [QQ\inv]_{kl} (DD_l + 2DR_l - RR_l) 
\end{eqnarray}
where $k$ now indexes the basis functions, $T$ refers to all of the data for the given tracer, and $QQ_{kl}$ is the product of basis function indexed by $k$ and $l$ for the random--random pairs.
\fi %%%%%%%%%%%%%%%%%%

In the case where the basis functions only depend on the pair separation, we have $f = f(|x_n - x_{n'}|)$ and $\xi = \xi(r)$, where $r$ is the separation at which to evaluate the correlation function.
Specifically, we can choose to define a set of basis functions $f^T$ as,
\begin{equation}
f^T_k(|x_n - x_n'|) =  i(g_k < |x_n - x_{n'}| < h_k)
\end{equation}
where $k$ denotes a particular bin in separation.
In this case the $DD$, $DR$ and $RR$ vectors become binned pair counts and the $QQ$ matrix becomes diagonal, with diagonal elements equal to the $RR$ vector elements. %TODO CONTINUE HERE
Then for this set of functions $f^T$, which is the definition of a top-hat (rectangular) function, the estimator is equivalent to that of Landy-Szalay.

\KSF{Worth giving analytic formulation of RR term? Or at least mentioning that there is one and we have derived it?}

\KSF{Discuss the limit of infinitesimal bins? Do we know what this is?}

We implement this estimator based on the correlation function package \texttt{Corrfunc} by \cite{Sinha2019}.
The code is open-source and available at \texttt{github.com/kstoreyf/Corrfunc}.
\KSF{Should give any detail about implementation?}

\KSF{where to put this link? In a footnote? And is it ok to keep it as a fork?}

\section{Applications}
\label{sec:app}

\KSF{Does this count as Applications if its on fake data? Better name?}

We demonstrate the application of \est on artifical data.
We generate lognormal mock catalogs \citep{ColesJones1991} using the \texttt{nbodykit} package \citep{Hand2018}.
We use an input power spectrum with the Planck cosmology \KSF{CITE}.
The true correlation function is then known: it is the Fourier transform of the input power spectrum, computed numerically.
Our test catalogs have size (750 \hmpc)$^3$ and a galaxy number density of $3 \times 10^{-4}$, to match that of the Brightest Cluster Galaxies \KSF{CITE}.
We construct 100 realizations of this box. 
We also generate a uniformly sampled random catalog with 10 times the number of galaxies as in the data catalogs.

\subsection{Demonstration using Spline Basis Functions}
\label{sec:spline}

\label{fig:spline}
\begin{figure}[ht]
\centering
    \includegraphics[width=0.8\textwidth]{tophat_cubic}
    \caption{A comparison between \est with a cubic spline basis function (red dotted) and the standard tophat estimator (blue solid). The shaded region is the 1$\sigma$ variation in the 100 mock catalogs. The cubic spline estimator has lower bias and lower variance with fewer bins. \KSF{quote rmse in caption? otherwise not clear from figure its less biased.} \KSF{does tophat need to be a different linestyle?}} \KSF{fix to dotted linestyle}

\end{figure}

\KSF{Should i have a figure showing the basis functions?}

\KSF{These sims have the weird thing where the data says the peak is offset. Should figure out, and also it's distracting from the main point of the figure}

\KSF{FINAL PARAMS:  number density 3e-4 (now 1e-4). rmax 200 hMpc (now 150). rcont=1000 (now rcont=300). L=1000 hMpc (now 750 - need more to account for higher rmax)}

A natural extension of tophat basis functions is the B-spline.
B-splines of order $n$ are piecewise polynomials of order $n-1$; they constitute the basis functions for spline interpolation \KSF{CITE}.
They have the nice property that the functions and their derivatives can be continuous, depending on the order.
Further, B-splines are well-localized, which provides a more direct comparison to the typical tophat basis (which is entirely localized).
For this demonstration we use fourth-order B-splines, which constitute the set of basis functions for a cubic spline, as they are the lowest-order spline to have a continuous first derivative.
We compare this standard estimator, reformulated as continuous functions using a tophat basis.
For the tophat basis we use 44 basis functions in the range $40 < r < 150$ \hmpc, each with a width of 2.5 \hmpc. 
For the cubic spline basis, we use the same $r$ range, but with only 11 basis functions, and knots chosen on a grid of 10 \hmpc.
\KSF{need to specify more about knots? is this accurate (not quite, the knots repeat values at the edge i think...) and should i mention control points?}

The results are shown in Figure~\ref{fig:spline}, compared to the true input \cf.
The spline basis results in a \cf estimate that has a root mean square error (RMSE) with respect to the truth of $4.38 \times 10^{-4}$, compared to $5.31 \times 10^{-4}$ for the tophat basis.
This holds true when we average over the bin \KSF{need to do weighted average based on which r-values contribute? supposed to hear from tinker}, as in standard practice, and then compute the RMSE.
The spline basis also results in lower variance across the 100 mocks, compared to the tophat basis.
Thus, \est allows for a more accurate and precise estimate of the \cf with fewer components.

We note that this is fundamentally different than a kernel-based estimator.
In a kernel formulation, each data point is smoothed by a particular kernel, smoothing out features in the resulting function.
With our estimator, the basis functions are fixed and the data is projected onto them.
This preserves the information in the data to the degree given by the chosen set of basis functions, which can in fact enhance features rather than smooth them.
\KSF{please add/correct here}

\subsection{BAO Scale Estimation Test}

Measurement of the BAO scale provides a good use case for our estimator.
The BAO feature is a peak in clustering on large scales, $\sim$150 Mpc, making it less sensitive to small-scale astrophysical effects.
It is one of the best tools for constraining cosmological models, in particular the distance-redshift relation \citep{Kazin2010, Anderson2012, Anderson2014, Alam2016}.

\KSF{Is there meaning to ``redshift space separation'' when I'm workin in a simulation box? Should I just say separation?}

\KSF{Should this be done on the periodic box? On a box assuming non-periodicity so we need the RR term? On a sky projection of a box, but without a window function?}

\KSF{what else do I need to cite for BAO?}

We base our BAO estimation on the method of the BOSS DR10 and 11 analysis \citep{Anderson2014}.
We measure the spherically averaged redshift space correlation function, $\xi(s)$, where $s$ is the redshift-space separation between pairs.
The BAO peak position is a function of multiple distance scales.
We can measure it by comparing the data to a fiducial model of clustering.
To a very good approximation, the peak position depends on the single scale dilation parameter
\begin{equation}
\alpha = \Bigg( \frac{D_A(z)}{D_A^{mod}(z)} \Bigg)^{2/3} \Bigg( \frac{H^{mod}(z)}{H(z)} \Bigg)^{1/3} \Bigg( \frac{r_s^{mod}}{r_s} \Bigg),
\end{equation}
where $D_A$ is the angular diameter distance, $H$ is the Hubble constant, $r_s$ is the sound horizon scale at the drag epoch, and the superscript ``$mod$'' denotes the value for the fiducial model. \KSF{mod or fid? fid just was v close to fit}
If we find $\alpha>1$, the peak is shifted to smaller scales than its position in the model, and if $\alpha<1$, the peak is shifted to larger scales.

% Using spherically averaged so don't need this!!
% The multipoles of the correlation function are given by
% \begin{equation}
% \xi_l(s) = \frac{1}{2} \int^{1}_{-1} \xi_l(s, \mu) L_l(\mu) d\mu
% \end{equation}
% where $L_l$ is the Legendre multipole of order $l$,  $s$ is the redshift-space separation between pairs, and $\mu = cos(\theta)$, $\theta$ is the angle between the separation vector $s$ and the line-of-sight direction (i.e. $\mu=1$ indicates a galaxy pair along the line of sight, $\mu$=0 indicates a pair with only transverse separation).

In standard practice, a fitting function is used to determine the value of $\alpha$:
\begin{equation}
\xi^{fit}(s) = B^2 \xi^{mod}(\alpha s) + \frac{a_1}{s^2} + \frac{a_2}{s} + a_3
\end{equation}
where $B$ is a constant that allows for a large-scale bias, and $a_1$, $a_2$, and $a_3$ are nuisance parameters to account for the broadband shape.
The model is given by \KSF{DESCRIBE}, and a $\chi^2$ fit is performed with five free parameters: $\alpha$, $B$, $a_1$, $a_2$, and $a_3$.
The resulting value for $\alpha$ is used to derive the actual values of the distance scales of interest.
Typically, density-field reconstruction is performed before applying the estimator to correct for nonlinear growth around the BAO scale \citep{Eisenstein2007}; for our toy example, we will omit this step.

The form of the standard fitting function is well-suited to our estimator, as it is a few-parameter model with a linear combination of terms.
To use our estimator to estimate $\alpha$, we take the numerical partial derivative of the model with respect to $\alpha$, using a change in $\alpha$ of size $d\alpha$.
Our fitting function is then
\KSF{Check if the derivative part is right!}
\begin{equation}
\xi^{fit}(s) = B^2 \xi^{mod}(s) + C d\alpha\frac{d\xi_{mod}(\alpha s)}{d\alpha} + \frac{a_1}{s^2} + \frac{a_2}{s} + a_3,
\end{equation}
where $C$ describes the contribution of $\alpha$.
That is, a value of $C=0$ indicates that the model is the best fit to the data and no scale dilation is needed.
More generally, we can use $C$ to compute $\alpha$ with $\alpha = 1 + Cd\alpha$.
With the terms in this fitting function the bases of our estimator, the values $B$, $C$, $a_1$, $a_2$, and $a_3$ are then given by the output amplitude vector $a$ as described in $\S$\ref{sec:est}.

\KSF{Choose realistic alpha values}

\KSF{Instead of this alpha test, should do a slightly more realistic one where we change the model cosmology? Issue is that we then don't know the truth}

\KSF{compare to tophat fit with standard BAO fitting function??}

\label{fig:bao_bases}
\begin{figure}[ht]
\centering
    \includegraphics[width=0.8\textwidth]{bao_bases}
    \caption{The set of basis functions used to fit for the BAO scale using our estimator. The $B^2$ term (green) is the fiducial model used to determine the scale dilation parameter $\alpha$. The $C$ term is the derivative of this model with respect to $\alpha$, allowing for the estimation of this parameter. The $a_1$, $a_2$, and $a_3$ terms are nuisance parameters to fit the broadband shape. \KSF{Make colorblind friendly! Do i also need to use different linestyles?}}
\end{figure}

We demonstrate this method using the same set of lognormal mocks as in $\S$\ref{sec:spline}.
We construct a simple test in which our model is the known input model, the Planck parameters, but with some scale dilation offset, here $\alpha_{mod}=0.98$ (while the true value is, of course, $\alpha=1$).
The set of basis functions for our estimator is shown in Figure~\ref{fig:bao_bases}.

We perform an iterative procedure to estimate $\alpha$.
After the first estimation with the $\alpha=0.98$, model, we use the value of $C$ to compute the recovered $\alpha$, as described above.
We then take that value as our $\alpha_{mod}$, and re-run the estimator.
We stop when the percent change between $\alpha_{mod}$ and the recovered $\alpha$ is less than 0.01\%; it typically takes only $\sim$5 iterations to converge. \KSF{update this with true values}
We apply this procedure to each of the 100 mocks; the resulting estimate for the correlation function is shown in Figure~\ref{fig:bao}.
The mean BAO estimate is shown in orange, and the mean tophat estimate is in blue; the truth is in black.
Our estimator clearly better represents the shape of the known \cf.
The mean value of the final recovered scale dilation parameter is $\alpha=0.9995$ \KSF{What do i quote for the error here?}, very close \KSF{quote percent?} to the true value of $1$.

\label{fig:bao}
\begin{figure}[th]
\centering
    \includegraphics[width=0.8\textwidth]{bao}
    \caption{Estimation of the correlation function using our estimator with basis functions based on the BAO fitting function (orange dot-dashed). The line is the mean of the final estimate from the iteration procedure for 100 mocks, and the shaded region is the $1\sigma$ variation. We also show the standard Landy-Szalay estimator, displayed as a tophat function (blue), as well as the true input correlation function (black). \KSF{This is for the 1e-4 box, need to run bigger box.} \KSF{Should this and the basis set figure be a single figure with two panels?} \KSF{Should i have a figure showing the sum of the basis functions and how that gets you the final cf?}}
\end{figure}

We note that these basis functions are significantly different than the tophat or B-spline bases previously explored, mainly because they are not localized.
This means that data at all scales could contribute to all basis functions.
It is then critical to ensure that the final parameter estimate does not rely on the range of scales chosen.
We have confirmed that in this application, the result is robust to the chosen range as long as the scales cover the range $40<r<200$ \hmpc, the typical range used in BAO analyses. \KSF{Actually have to check this and update numbers!}

% \section{Application to survey data} \label{sec:app}

% We apply the estimator to the SDSS DR3 sample from \cite{Eisenstein2005}.

% The reconstructed dataset we use is described in \cite{Kazin2010}.

\section{Discussion} \label{sec:discuss}

\subsection{Beyond the Landy-Szalay Estimator}

While we have formulated our estimator as a generalization of \LS, as it is the standard used in \cf analyses and has optimal properties under certain conditions, we can also reformulate it for other estimators.
The formulation currently requires a normalization term (i.e. denominator) of the $RR$ counts, as we replace this with our $QQ$ term.
This is the case for the \cite{PeeblesHauser1974} estimator and the \cite{Hewett1982} estimator:
\begin{eqnarray}
    \xi_{PH}(r) &=& \frac{DD - RR}{RR} \\
    \xi_{Hew}(r) &=& \frac{DD - DR}{RR}.
\end{eqnarray}
\KSF{dropped $k$ and $(r)$ notation here, do i need?}
We could also generalize estimators which have a $DR$ term as the denominator, such as the \cite{DavisPeebles1983} estimator,
\begin{equation}
    \xi_{DP}(r) = \frac{DD - DR}{DR}
\end{equation}
by defining
\begin{equation}
    DQ = \frac{2}{N_D N_R} \sum_{n m} f(T_n, T_m) \cdot f\T(T_n, T_m).
\end{equation}
This could be extended to almost any linear combination of pair counts.
The estimator of \cite{VargasMagana2013} selects the optimal combination of pair counts; our estimators could be combined to create an even more generalized estimator.
\KSF{some of the terms in V-M include DD in the denom - need to mention this?}

\KSF{Mention other work with Barnett about LS estimator?}

\subsection{Computational Performance}

The computational scaling is by definition the same compared to traditional estimators due to the limiting factor of pair-finding.
\Est has the additional need for evaluating the function $f$ for each pair of galaxies.
For simple basis functions like splines, this will only marginally decrease performance.
For more complicated functions such as evaluating a cosmological model, \est may incur extra computational expense.
Basis functions can also be input on a grid and then interpolated; the performance is then the same for all functions, but the interpolation for each function for each pair does somewhat decrease the performance.

\KSF{Is there a need to say much more? If not, maybe doesn't deserve its own section - could this go in the implementation section? Or maybe tacked onto another small section if we have one that makes sense}

\subsection{Effect on Covariance Matrix Estimation}

We have shown that \est results in \cf estimates that are just as accurate with fewer components.
This is critical when estimating the covariance matrix, which is necessary for parameter inference.
The covariance matrix is difficult to compute theoretically;instead, it is usually estimated by evaluating the \cf on a large number of mock catalogs and computing the covariance between the bins (e.g. \citealt{Reid2010}; \citealt{Anderson2014}).
The unbiased estimator for the sample covariance matrix is (e.g. \citealt{Anderson2003})
\begin{equation}
\hat{C}^{ML}_{ij} = \frac{1}{N_{mocks}-1} \sum_{k=1}^{N_{mocks}} \bigg( x^k_i - \bar{x}_i \bigg) \bigg( x^k_j - \bar{x}_j \bigg),
\end{equation}
where $k$ denotes the index of the mock, $i$ and $j$ denote the index of the bin or component, $x$ denotes the estimate in that bin for that mock, and $\bar{x}$ denotes the mean value of the estimate in that bin across the mocks.
To get an unbiased estimate of the inverse covariance matrix, we require a correction factor, as the inverse of an unbiased estimator is not necessarily unbiased.
The unbiased estimator for the sample inverse covariance matrix can be shown to be \citep{Hartlap2007}
\begin{equation}
\hat{C}\inv = \frac{N_{mocks}-N_{bins}-2}{N_{mocks}-1} \bigg( \hat{C}^{ML}\bigg) \inv.
\end{equation}

%This prefactor correction results in the  propagates to the variance in 
The variance in the elements of this estimator then have a dependence on $N_{mocks}$ and $N_{bins}$.
This propagates to the derived cosmological parameters, resulting in an overestimation of the error bars (\citealt{Hartlap2007}; \citealt{Dodelson2013} \citealt{Percival2014}; \citealt{TaylorJoachimi2014}).
Assuming that $N_{mocks} >> N_{bins}$ (and both much larger than the number of parameters to be estimated), and that the measurements are Gaussian distribued, the error bars are inflated by a factor of $(1 + N_{bins}/N_{mocks})$ (i.e., the true constraints are tighter than the derived ones).
This factor becomes critical at the precision of cosmological parameter estimation \citep{Percival2014}.

Typically, this is dealt with by generating a very large number of mocks.
For the Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{Dawson2013}), $\sim$600 mocks were needed and the analysis used 41 bins \citep{Sanchez2012}.
Future surveys will have more costly requirements on mock catalogs, with larger simulations necessary to cover the larger survey volumes.

An alternative to increasing $N_{mocks}$ is decreasing $N_{bins}$ to achieve the same error on precision.
In the standard method, this is shown to \emph{increase} the statistical error, albeit only slightly \cite{Percival2014}.
A substantial increase in bin width would prevent capturing information in finer clustering features; even the relatively broad BAO peak requires a bin size on the order of its width of $\sim$10\hmpc.
In fact, in the standard method more bins would be desireable, but the number is limited by the available number of mocks due to the error discussed above.

With our estimator, we have shown that we can reduce the variance by using fewer components, without sacrificing accuracy.
This means that we can safely reduce $N_{bins}$, or in our replacement of bins with continuous functions, the number of basis functions $K$.
The covariance matrix will be the covariance between these basis functions. \KSF{worth noting that the structure of this covmat will be significantly different, esp if non-orthogonal?}
To then achieve the same precision on the error on the cosmological parameters, a lower value of $N_{mocks}$ becomes possible.
This will significantly reduce requirements on mocks, which will be particularly important for upcoming large surveys.

\KSF{I think the result of discussions was that there wasn't a good way of showing this without propagating all the way to cosmological parameters. Would love a figure showing lower covariance errors but not sure how without full propagation}

\subsection{Further Applications}

The formulation of \est opens up many possibilities for extracting information from the correlation function.
The most straightforward applications are standard basis functions or linearizeable astrophysical models, as we have shown here.
Other applications for the direct estimation of cosmological parameters could include the growth rate of cosmic structure $f$ \citep{Satpathy2016, Reid2018} and primordial non-Gaussianity in the local density field $f^{local}_{NL}$ \citep{Karagiannis2014}.
\KSF{mention idea of doing full cosmo model analysis by taking derivs wrt cosmological params? cool but less connected to citeable papers perhaps}

We can take our estimator a step further by choosing basis functions that depend not only on the separation between tracer pairs, but also on the properties of the tracers themselves.
One such application is the redshift dependence of the Alcock-Paczynski effect \cite{AlcockPaczynski1979}, which can be used to constrain the matter density $\Omega_m$ and the dark energy equation of state parameter $w$ \citep{Li2016}.
The basis functions $f$ in this case would take the form
\begin{equation}
    f_k(T_n, T_m) = f_k(r_{nm}, z_n, z_m),
\end{equation}
where $z$ is the redshift of tracer $n$ or $m$.
Another potential use case is the luminosity and color dependence of galaxy clustering, which can be used to probe galaxy formation \citep{Zehavi2011}.
This could be extended to other galaxy properties.

The estimator gives us the opportunity to investigate more subtle or exotic signals which are anomalous with respect to our conventional models.
Anomalies could appear as inhomogeneities or anisotropies in the data.
For example, cosmological parameters could vary across the sky, which has previously investigated in patches across the Cosmic Microwave Background \citep{MukherjeeWandelt2018}.
Another possibility is anisotropy in the cosmic acceleration, which could appear in various probes including the BAO peak \citep{Faltenbacher2012} and Type Ia supernovae \citep{Colin2019}.
With our estimator, we could introduce a dependence on location or direction into our basis functions, and constrain the potential deviation from homogeneity or isotropy.
While these effects would be highly degenerate with systematics, our estimator combined with robust systematics mitigation allows us to investigate the possibility of new physics.

Finally, our estimator can be directly related to a power spectrum analysis.
We could use a Fourier basis as our set of continuous functions.
This would allow us to directly project the data onto Fourier modes.
This represents a step towards unifying the correlation function and the power spectrum. \KSF{there's more to say here but I'm not sure what}

\section{Summary}

We did cool science!

\KSF{Do we want a summary? Think it might be useful for this particular paper, but def repetitive}

\acknowledgements
It is a pleasure to thank\ldots

\appendix
\section{Affine Invariance}\label{sec:affine}

The estimator is invariant under an affine transformation. We represent this by a matrix M, such that 
\begin{equation}
f' \leftarrow Mf
\end{equation}
Then in the primed basis, the pair counts become
\begin{eqnarray}\displaystyle
DD' &=& \sum_{n n'} M f_{n n'} = M\,DD
\\
DR' &=& \sum_{n m} M f_{n m} = M\,DR
\\
RR' &=& \sum_{m m'} M f_{m m'} = M\,RR
\end{eqnarray}
where $f_{n n'} = f_k(T_n, T_{n'})$.
We have factored $M$ out of the summation. For the $QQ$ matrix we have
\begin{eqnarray}\displaystyle
QQ' &=& \sum_{m m'} (Mf_{m m'}) \cdot (Mf_{m m'})\T \\
&=& M\Bigg[ \sum_{m m'} f_{m m'} \cdot f_{m m'}\T \Bigg]M\T \\
&=& M QQ M\T
\end{eqnarray}
Then the amplitudes in the primed basis become
\begin{eqnarray}\displaystyle
a' &=& [M QQ M\T]\inv \cdot [MDD - 2\,MDR + MRR] \\
&=& (M\T)\inv QQ\inv \cdot [DD - 2\,DR + RR] \\
&=& (M\T)\inv a
\end{eqnarray}
and the estimator in the primed basis is 
\begin{eqnarray}\displaystyle
\xi' &=& [(M\T)\inv a]\T \cdot (Mf) \\
&=& a\T[(M\inv)\T]\T \cdot (Mf) \\
&=& a\T M\inv \cdot Mf \\
&=& a\T \cdot f = \xi.
\end{eqnarray}
Thus after an affine transformation of the basis function, the resulting estimator is equivalent to the estimator in the original basis.

We note that this requires $M$ be invertible.
% not sure if i understand what i wrote here
However, any two equivalent bases must be related by the inverse of a transformation matrix, so this requirement is already satisfied.

%need style file error
%\bibliographystyle{apj} 
%\bibliography{paper}
% To copy from mendeley locally: in paper dir, [cp ~/code/bibtex/LSS-est_paper.bib .]
\bibliography{LSS-est_paper}

\end{document}